What can GitHub and Googlecode Issues tell us about Exceptions in Android Apps?

### 2014-08-22 android-stacks.tex
Changing the focus of the paper to android only.

\begin{document}
\title{ What can Stack Traces tell us about Java Exceptions? }

%\title{A first Glance of what is causing Mobile Crashes: A GitHub and Googlecode Repository Mining Study }

@@

In this work, we perform the first large scale analysis of Java stack traces and
pinpoint how they can reveal bad programming practices. We mined and analyzed
the stack traces embedded in all issues of Java projects available GitHub.
Overall, our research set includes 356,057 issues defined at 16,836 projects, 
from which 28,800 stack traces were extracted. From this set, the stack traces
of 482 Android projects were investigated in more detail, in combination with
source code and bytecode analysis. In this study some patterns of exception
(mis)-use were consistently detected such as: unexpected wrappings (e.g., Errors
being wrapped in checked exceptions), revealing that Java hybrid exception
model is not fully used according to its purpose;  undocumented runtime
exceptions signaled by third party code, which is a serious threat to program
dependability; and a high prevalence of java.lang exceptions reported on issues,  
they were found in aproximately 50\% of all analyzed exception stack traces.

@@

Modern applications have to cope with an increasing number of abnormal
computation states that arise as a consequence of faults in the application
itself (e.g., access of null references), noisy user inputs or faults in
underlying middleware or hardware. The exception handling
mechanism~\cite{goodenough1975exception} is one of the most used schemes for
detecting and recovering from such exceptional conditions. When an exception is
thrown it can be caught inside the application (by the program's exception
handling code) and the application has the opportunity to assess the conditions
that caused the fault and either recover from them or abort. If the exception
that remains uncaught, the program crashes. Uncaught exceptions are one of the
main causes of application crashes~\cite{jo2004uncaught}.

In most programming languages featuring exception handling, when a crash
happens, the language runtime generates debugging information to help the
programmer identify the cause of the crash. The debugging information is
typically a stack trace, which includes an explanatory message and the execution
stack frame when the exception occurred.
%In Java, stack trace consists of an ordered list of methods that were active on
%the call stack before the exception occurred.

Although the exception handling mechanism has been the target of several studies
(e.g.~\cite{miller1997issues,Robil00,shah2010understanding,
garcia2007extracting,garcia2001comparative,cabral2007exception,coelho2011unveiling}),
the exception handling code is generally poorly understood and often the least
tested part of software systems~\cite{coelho2011unveiling}.  To analyze the
exception handling code, current studies use either static (source code-based)
analysis or dynamic (runtime-based) testing methods. Neither method can explore
the full execution space of a program and therefore they both fail to account
for exceptions implicitly signaled by the runtime environment (e.g. due to
resource depletion) or dynamically loaded code. Exception stack traces, when
available, enable the post-mortem analysis of exception handling behaviours of
programs and consequently the evaluation of exception handling mechanisms in
programming languages.

The availability of exception stack traces until recently was scarce, while the
available data concerned a few large scale projects (e.g. the Mozilla
crash report
dataset\footnote{https://code.google.com/p/promisedata/issues/detail?id=88}).
Along the recent years, we have witnessed a rapid increase in the number of open
source projects that make their control version and issue tracking systems
publicly available in repository hosting sites such as GitHub. The
concentration of open source projects and the existence of initiatives that mine
such information, such as GHTorrent~\cite{Gousi13}, allow the information about
projects to be visible within and across projects and make a full population
studies possible.

In this work, we perform the first post mortem analysis of exception handling in
Java. Using a custom tool called ExceptionMiner, which we develop specifically
for this study, we mine stack traces from issues across all Java projects on
GitHub. Overall, we analyzed 356,057 issues from 16,836 projects, from which
28,800 stack traces were extracted and used as a source of information for our
analysis. The stack trace analysis was augmented by additional information
extracted using byte code and source code analysis done on a carefully selected
subset of our initial sample (482 Android projects). Manual inspection was
used to leverage the understanding of stack traces and support
further discussions and insights.
To guide our analysis, we compiled a set of best practices on how to use
exceptions by composing relevant suggestions by Gosling~\cite{gosling2000java},
Wirfs-Brock~\cite{wirfs2006toward} and Bloch~\cite{bloch2008effective}.  The
comparison of those with the mined exceptions revealed several interesting
findings, such as:

@@

  \item  Odd exception chains (e.g., Error wrapping checked and runtime
    exceptions and vice-versa), indicating that the purpose of Java's hybrid
    exception model may not have been adequately used.

  \item  Issues reporting uncaught checked exceptions.

@@

The contributions of this study are as follows:
\begin{itemize}

  \item  It performs the first large scale analysis of Java stack traces and how
    they can reveal bad-practices on the use of exceptions in Java.

@@

@@

These contributions are relevant to:
(i) developers of Java applications
to familiarize with the most common reported exceptions mis-uses which is the
first step to help developers to avoid making them;
(ii) designers of languages
to consider ways of reducing the abundance of NullPointerExceptions;
and (iii)
tool designers to consider developing tools that enable a cross-project defect
analysis.

@@

%This paper is organized as follows. Section 2 presents a
%background on exception handling mechanism. Section 3 presents the study design.
%Section 4 reports study findings. Section 5 provides further discussions and insights.
%Section 6 presents the threats to validity associated to this study. Finally Section
%7 describes the related work, and Section 8 presents our conclusions and
%directions for future work.

@@

\subsection{Checked an Unchecked Exceptions in Java} \label{sec:extypes}

In Java, an exception is represented according to the class hierarchy shown in
Figure~\ref{fig:exchier}.  According to it every exception is an
instance of the Throwable class, and can be of three kinds: checked exceptions
(extends Exception), run-time exceptions (extends RuntimeException), and errors
(extends Error)~\cite{gosling2000java}. Neither run-time exceptions nor errors need to be specified on method
signatures, and hence both are referred as unchecked exceptions.

@@

\begin{figure} \centering \includegraphics[scale=0.6]{wrappings}
\caption{Example of an Exception stack trace in Java.}
\label{fig:wrapping}
\end{figure}

@@

  \item[RQ2] What patterns of exception (mis)-use emerge from the exception stack trace
    analysis?

@@

@@

Based on the assumption that regardless the issue type, every exception stack
trace contains relevant information concerning the exception structure of the
projects analyzed, we opted for not restricting the analysis only on defect
issues.\footnote{We conducted the same analysis on the defect issues and the top
exceptions are similar to the ones mentioned on defect issues. Due to space
limitation we limit to present the general analysis here. More detailed analysis
on the defect issues can be found at:
\url{www.dimap.ufrn.br/~roberta/icsme2014}}

@@

After distilling the information available on the 28,800 stack traces (extracted from the 
issues and issues comments of 16,836 Java projects) we found 607 distinct exceptions
reported as the root causes of stack traces.

Table \ref{tab:toptenjava} illustrates the top 10 root couses found in the study ranked by the number of distinct
project repositories on which they were reported. The table illustrates the
number of times each exception was reported as the root cause of a stack trace
, the number of distinct projects reporting such causes on stack traces, and the percentages of
such in relation to all reported exceptions and analyzed projects respectaly.

\begin{table}
\begin{tabular}{lrrrr}
    \hline
    \bfseries{Exception} &  \multicolumn{2}{c}{\bfseries{Occurrences}} &  \multicolumn{2}{c}{\bfseries{Projects}}\\
    & \bfseries{\#} &  \bfseries{\%} & \bfseries{\# } & \bfseries{\% } \\
    \hline
java.lang.NullPointerException             & 7578 & 26,31\% & 1671 & 44,47 \\
java.lang.IllegalArgumentException         & 1573 & 5,46\%  & 670  & 17,83\% \\
java.lang.ClassNotFoundException           & 1203 & 4,18\%  & 514  & 13,68\% \\
java.lang.RuntimeException                 & 1093 & 3,80\%  & 411  & 10,94\% \\
java.lang.IllegalStateException            & 964  & 3,35\%  & 431  & 11,47\% \\
java.lang.ClassCastException               & 886  & 3,08\%  & 408  & 10,86\% \\
java.lang.ArrayIndexOutOfBoundsException   & 834  & 2,90\%  & 345  & 9,18\% \\
java.lang.NoSuchMethodError                & 822  & 2,85\%  & 369  & 9,82\% \\
java.io.IOException                        & 656  & 2,28\%  & 337  & 8,97\% \\
java.lang.OutOfMemoryError                 & 601  & 2,09\%  & 235  & 6,25\% \\
%java.lang.NoClassDefFoundError            & 554  & 1,92\%  & 299  & 7,96\% \\
%java.util.ConcurrentModificationException & 465  & 1,61\%  & 180  & 4,79\% \\
%java.lang.IndexOutOfBoundsException       & 445  & 1,55\%  & 205  & 5,46\% \\
\hline
All	 &	23956 &	100\%	& 3758 & 100\%	 \\
\hline
  \end{tabular}
\caption{Top 10 most popular exceptions on issues in Java projects}
\label{tab:toptenjava}
\end{table}

@@

@@

\begin{table}
  \centering
  \begin{tabular}{|p{2cm}| p{5cm}|}
    \hline
    \bfseries{Category} & \bfseries{Description} \\
    \hline
      Programming logic &  Exceptions implicitly  thrown by the JVM when the 
evaluation of an expression violates the normal semantics of language (e.g., 
out-of-bounds array index, division-by-zero, access to a null reference). 
Such exceptions are usually defined on java.lang and java.util packages. \\ \hline
      Resources (IO)                         & Exception releated to the manipulation of IO resources or device  (e.g., file, network) \\ \hline
      Security                               & Exception releated to security issues (e.g. password validation, criptography) \\ \hline
      Concurrency                            & Exceptions related to multi-threaded programming \\ \hline
      Backward compatibility                 &  A list of exceptions related to backward compatibility - a complete list can be found here [ref]  \\ \hline
      Reflection                             & Exceptions related to Java reflection library  \\ \hline
      Specific                & Exceptions created for specific applications, frameworks or libraries \\ \hline
      General (Exception, Error, Runtime)    & The top level exceptions in Java    \\ \hline
    \hline
  \end{tabular}
  \caption{Description of categories.}
  \label{tab:categories}
\end{table}

@@

\begin{table}
  \centering
  \begin{tabular}{lrr}
    \hline
    \bfseries{Category} & \bfseries{Java} & \bfseries{Android Subset} \\
    \hline
      Programming logic (java.lang and util) & 12625  (41,10\%) & 2235 (55,64\%)\\
      Resources (IO)                         & 12440 (40,50\%)  & 727 (18,10\%) \\
      Security                               & 220  (0,72\%)    & 165 (4,11\%)\\
      Concurrency                            & 633 (2,06\%)     & 116  (2,89\%)\\
      Backward compatibility                 & 1580 (5,14\%)    & 219 (5,45\%) \\
      Reflection                             & 1413 (4,60\%)    & 91 (2,27\%)\\
      Specific (GUI,FRAMEWORK)               & 340 (1,11\%)     & 197 (4,90\%)\\
      General (Error, Exception, Runtime)    & 1468 (4,78\%)    & 267 (6,65\%)\\
    \hline
  \end{tabular}
  \caption{Characterization of the 100 root causes.}
  \label{tab:tophundrend}
\end{table}

@@

As detailed in Section~\ref{sec:android}, 482 Android projects were carefully
selected to enable a deeper investigation of stack trace information, by
leveraging stack trace information with the source code and bytecode analysis.
For this subset of Android project, we could extract every package related to
each exception signaler, and based on this information we could classify every
stack trace based on its root signaler as presented in Table~\ref{tab:signalers}.

\begin{table}
  \centering
  \begin{tabular}{rp{29em}}
    \hline
    \bfseries{Signaler} & \bfseries{Description} \\
    \hline
    \bfseries{android} & If the exception is thrown by a method defined in Android Platform or OS, or in a JDK library used by them.\\
    \bfseries{app}     & If the exception is thrown by an application method or in a  DK library used by it.\\
    \bfseries{libcore} & If the exception is thrown by one of the core libraries reused by Android (i.e., org.apache.harmony, org.w3c.dom, sun.misc, org.apache.http, org.json, org.xml). \\
    \bfseries{lib}     & If the exception is thrown by a method that was not defined by any of the elements above.\\
    \bfseries{java}    & If all methods on the stack trace are JDK library methods.\\
    \hline
  \end{tabular}
  \caption{Sources of exceptions in Android}
  \label{tab:signalers}
\end{table}

For each exception we could discover the number of times it occurred on stacks,
the number of distinct projects on which such stacks were defined (and consequently its
popularity) and the of times it was signaled by different signaler types. We
could calculate the popularity of each root cause, considering the number of
distinct projects they were found and the whole set of projects analyzed with at
least one stack trace (482). Table~\ref{tab:toptenandroid} 
%and Figure~\ref{fig:androidsignaler}
presents the mined data.

\begin{table*}
  \centering
  \begin{tabular}{rccccccccc}
    \hline
    \bfseries{Root Exception} &  \multicolumn{2}{c}{\bfseries{Occurrences}} &  \multicolumn{2}{c}{\bfseries{Projects}} & \textsf{android} & \textsf{libcore} & \textsf{app} & \textsf{lib} & \textsf{java} \\
    & \bfseries{\#} &  \bfseries{\%} & \bfseries{\# } & \bfseries{\% } &&&&&\\
    \hline
java.lang.NullPointerException            & 1225 & 30,08\% & 254 & 52,70\% & 473 & 18 & 595 & 137 & 2 \\
java.lang.IllegalStateException           & 234  & 5,75\%  & 99  & 20,54\% & 165 & 12 & 36  & 20  & 1 \\
java.lang.IllegalArgumentException        & 255  & 6,26\%  & 94  & 19,50\% & 146 & 6  & 64  & 39  & 0 \\
java.lang.RuntimeException                & 232  & 5,70\%  & 91  & 18,88\% & 167 & 1  & 47  & 17  & 0 \\
java.lang.OutOfMemoryError                & 180  & 4,42\%  & 56  & 11,62\% & 121 & 15 & 17  & 23  & 4 \\
java.lang.NoClassDefFoundError            & 73   & 1,79\%  & 52  & 10,79\% & 9   & 0  & 37  & 26  & 1 \\
java.lang.ClassCastException              & 94   & 2,31\%  & 49  & 10,17\% & 43  & 0  & 40  & 11  & 0 \\
java.lang.IndexOutOfBoundsException       & 127  & 3,12\%  & 47  & 9,75\%  & 47  & 0  & 71  & 8   & 1 \\
java.lang.NoSuchMethodError               & 57   & 1,40\%  & 40  & 8,30\%  & 9   & 0  & 39  & 9   & 0 \\
java.util.ConcurrentModificationException & 54   & 1,33\%  & 38  & 7,88\%  & 5   & 0  & 43  & 6   & 0 \\

@@

To answer this question we analyzed general characteristics of reported stack traces and, in the case of stack traces reported on Android repositories, we combine stack trace information with
source code and byte code analyses to enable a more detailed analysis.
Doing so, we mined some characteristics of stacks that points to best practices violations.

\subsection{Undocumented runtime exceptions being thrown.}

To better investigate the extent of this problem we classified each stack trace according to the type of the
root exception. Table~\ref{tab:typeroottab} presents the types of root causes of all stack traces reported on 
the Android repositories. As we can see from Table ~\ref{tab:typeroottab},  most of the exceptions signaled by third-party code
 (e.g., Android and libs) were runtime exceptions. After filtering all the exceptions implicitly signaled by 
JVM (due to programming mistakes) and inspecting the signaler methods of such exceptions (i.e.,  direct an indirect 
instances of java.lang.RuntimeException). We could observe that only 1 method out of 118 inspected signalers
 (i.e., 0,8\%) documented the explicitly thrown runtime exception in a Javadoc comment; and none of these methods
included the runtime exception  (reported on the stack) as part of the exception interface of the method (i.e., using 
throws clause on method signature). This result is aligned with the results of other  study conducted by from 
Sacramento et al ~\cite{sacramento2006unchecked} which observed that the
runtime exceptions in .Net programs are most often not documented.

\begin{table}
\centering
\begin{tabular}{lcccccc}
    \hline
    \bfseries{Type} & \bfseries{Android} & \bfseries{Libcore} & \bfseries{App} & \bfseries{Lib} & \bfseries{Java} & \bfseries{All}\\
    \hline

Runtime	&	1075	&	56	&	1446	&	374	&	10	&	2961	\\
Error	&	144	&	34	&	168	&	105	&	9	&	460	\\
Checked	&	110	&	230	&	139	&	145	&	12	&	636	\\
Throwable	&	0	&	0	&	1	&	0	&	0	&	1	\\
Undefined	&	1	&	0	&	6	&	8	&	0	&	15	\\
    \hline
All		& 1330	&	320	&	1760	&	632	&	31	&	4073	\\
    \hline
  \end{tabular}
\caption{Types of root exceptions.}
  \label{tab:typeroottab}
\end{table}

\begin{figure}
\centering
%\includegraphics[width=\hsize]{chart_exceptiontypes.pdf}
\includegraphics[width=\hsize]{exception_types.pdf}
\caption{Distribution of Checked and Unchedked exceptions.}
\label{fig:typeroot}
\end{figure}

On the other hand, we could observe that most of the exceptions signaled by the set of 
libraries reused by Android platform (i.e., org.apache.harmony,
org.w3c.dom, sun.misc, org.apache.http, org.json, org.xml, javax -- referred in
this work as libcore) were checked exceptions. It is pointed as a good practice
for libraries (see Section~\ref{sec:best}) because by using checked exceptions
libraries can define a precise exception interface to its clients. Such
libraries are heavily used in several projects, and such precise exception
interface may be related to the library's maturity.

\subsection{Uncaught Checked Exceptions}

>>>>

\begin{document}
%\title{ What GitHub and Googlecode can tell us about the Causes of Crashes of Android Apps? }
%\title{ A Mining Study the Googlecode can tell us about the Causes of Crashes of Android Apps? }

\title{ What can GitHub and Googlecode Issues tell us about Exceptions in Android Apps? }

%\title{On the Common Characteristics Android-Apps Crashes: Results of a GitHub and Googlecode Mining Study}

%\title{On the Common Characteristics of Android-Apps Crashes: Results of a GitHub and Googlecode Mining Study}

%\title{On the Common Characteristics of Android Crashes: Results of a GitHub and Googlecode Mining Study}

%\title{A first Glance on Characteristics of Android Crashes: Results of a GitHub and Googlecode Mining Study}

@@

This paper reports a repository mining study which analyzed all the issues created on
 Android projects available in GitHub (482 applications) and Google code (141 appplications)
Overall, our research set includes XXXX issues from which XXX stack traces 
were extracted. Such stack traces were distilled and analyzed in 
combination with source code and bytecode analysis. In this study some patterns of exception
(mis)-use were consistently detected such as: a high prevalence of NullPointerExceptions and 
other java.lang exceptions reported on issues - they were found in aproximately 50\% of all analyzed 
exception stack traces; undocumented runtime exceptions signaled by third party code - which is a serious threat to program
dependability; unexpected wrappings (e.g., Errors being wrapped in checked exceptions), revealing that Java hybrid exception
model is not fully used according to its purpose.

@@

The pervasiveness of the World Wide Web and the astonishing increase in the
number of smartphones have stimulated the creation of a new class of
applications which extend the phones capabilities far beyond of the basic calls
and textual messages.

Such applications run over one or more platforms, among them the Android platform
leads the smartphone market, comprising around XXXX [GoogleIO2014], thanks to its
Java-based open-source model and hardware partners like Samsung, HTC, Motorola, Asus 
(besides Nexus its own device) [ref].

Such applications are intrinsically multi-threaded and the development model and used APIs 
can be  slightly or drastically different and more error prone than tradicional development [arie-icse].

Moreover, such applications have to cope with an increasing number of abnormal
computation states. Besides the known source of failures (i.e., programmming errors, noisy 
user inputs or faults in underlying middleware or hardware) such applications need to deal with abnormal states that 
 arise as a consequence of:  memory and batery restrictions [ref], faults on third party services and libraries, 
compatibility issues caused by constant platforms updates [ref], conection with devices such as GPS, camera.

Therefore, techniques for error detection and handling are not  an optional add-on but a fundamental part of any system~\cite{bruntink2006discovering}. The Android Platform uses the exception handling mechanism embeeded
 in Java to deal with such conditions. However, although the exception handling mechanism
~\cite{goodenough1975exception} is one of the most used schemes for
detecting and recovering from such exceptional conditions, studies (e.g.~\cite{miller1997issues,Robil00,shah2010understanding, garcia2007extracting,garcia2001comparative,cabral2007exception,coelho2011unveiling}),
have shown that it is generally poorly understood and the less tested parts of the system ~\cite{coelho2011unveiling}.
The consequence of the exception handling code may come the other way around making the code even less robust.

Our conjecture is that several Android applicatons may be suffering from similar vulnerabilities on the exception
handling code, and that such vulnerabiliries may be leading to application crashes.

Untill recent years, the availability of issues reporting exceptional conditions was scarce, while the
available data concerned a few large scale projects (e.g. the Mozilla
crash report dataset\footnote{https://code.google.com/p/promisedata/issues/detail?id=88}).
Along the recent years, we have witnessed a rapid increase in the number of open
source projects that make their control version and issue tracking systems
publicly available in repository hosting sites such as GitHub and Googlecode.

% The
%concentration of open source projects and the existence of initiatives that mine
%such information, such as GHTorrent~\cite{Gousi13}, allow the information about
%projects to be visible within and across projects and make a full population
%studies possible.

Among the issues defined on issue trackers, we can find issues reporting bugs, and system crashes 
found as a result of application executions. Such issues usually contain exception stack traces [ref][ref]. 
When available, such information can enable different kinds of post-mortem analysis, such as:
: classify .... (get from previous versions).

Such analysis focuses a system at a time, and none of them target a family of Android Application
(which share a common core) looking for similar characteristics and vulnerabilitys on the crashes reported.

In this work, we perform the first post mortem analysis of the exception stack traces embedded on issues
of all Android applications available in GitHub and Googlecode. Using a custom tool called ExceptionMiner,
 which we develop specifically for this study, we mine stack traces from issues across all Android projects on
GitHub and Googlecode. Overall, we analyzed the XXXX issues extracted from on XXX Android projects defined in
GitHub and XXX projects defined in Googlecode. From these issues  XXX stack traces were extracted and used as
 a source of information for our analysis. The stack trace analysis was augmented by additional information
extracted from JDK, Android SO and Android SDK using byte code and source code analysis - which was done 
on a carefully selected subset of our initial sample (482 Android projects). Manual inspection was
used to leverage the understanding of exception stack traces and support further discussions and insights.

%To guide our analysis, we compiled a set of best practices on how to use
%exceptions by composing relevant suggestions by Gosling~\cite{gosling2000java},
%Wirfs-Brock~\cite{wirfs2006toward} and Bloch~\cite{bloch2008effective}.  The
%comparison of those with the mined exceptions revealed several interesting
%findings, such as:

Some outcomes consistently detected during this study were the following:

@@

   \item  Unexpected exception wrappings (e.g., Error wrapping checked and runtime
    exceptions and vice-versa), indicating that the purpose of Java's hybrid exception model may not have been adequately used.

   \item  The development model of Android platform (strongly based on system call-backs) encouraged 
     the use of general RuntimeExceptions.

  %\item  Issues reporting uncaught checked exceptions.

@@

Although the study focused on Android opensource applications, some of the results are not restricted to this context, 
for instance: some of the unexpected wrappings were performed by Java libraries and elements (e.g. reflection library and 
the static blocks); and the undocumented runtime exceptions were also found in Java libraries beyonds the Android plaftorm.
The study results are relevant to: (i) Android and Java developers to familiarize with the most common reported exceptions 
mis-uses which is the first step to help developers to avoid making them; (ii) designers of languages to consider ways of reducing the abundance of NullPointerExceptions; and (iii) tool designers to consider developing tools that enable a cross-project defect analysis.

Hence, the main contributions of this study are as follows:
\begin{itemize}

  \item  It performs a large scale post morten analysis of exception stack traces of Android projects.

@@

  \item  It presents the the exploratory study methodology to ..... (complete - view the reviewer comment)

@@

@@

The ramaining of this paper is organized as follows. Section 2 presents a
background on Android Plaform and Java exception handling mechanism. 
Section 3 presents the study design. Section 4 reports study findings. 
Section 5 provides further discussions and insights.
Section 6 presents the threats to validity associated to this study. Finally Section
7 describes the related work, and Section 8 presents our conclusions and
directions for future work.

@@

\subsection{Java Exception Handling Mechanism} \label{sec:extypes}

In Java, an exception is represented according to the class hierarchy shown in
Figure~\ref{fig:exchier}.  According to it every exception is an
instance of the Throwable class, and can be of three kinds: the checked exceptions
(extends Exception) - means that the compiler statically checks if appropriate handlers 
are provided within the system, to enable such checking each checked exception must be specified 
on the method signature;  the run-time exceptions (extends RuntimeException) and errors
(extends Error)~\cite{gosling2000java}  -  which do not need to be specified on method
signatures, and hence both are referred as unchecked exceptions.

@@

\begin{figure} \centering \includegraphics[scale=0.]{stacktrace.png}
\caption{Example of an Exception stack trace in Java.}
\label{fig:wrapping}
\end{figure}

% \subsection{Public Project Repositories and Issue Trackers} 
% ONE OF THE REVIEWERS COMPLAINED ABOUT THE ABSENCE OF EXPLANATIONS OF WHAT AN ISSU IS...
% As mentioned before, currently more and more projects are becomming available in 
% public project repositories such as GitHub and Google Code.
% Such repositories provides facilities such as....
% The issue trackers available in such projects ....

\subsection{The Android Platform} \label{sec:extypes}

\begin{figure} \centering \includegraphics[scale=0.8]{arch3.png}
  \caption{Android Archtecture} \label{fig:exchier} \end{figure}

@@

  \item[RQ2] What patterns of exception (mis)-use related to exception types and exception propagation emerge from the exception stack trace analysis?

@@

\subsubsection{Google Code}

Until January 2014 there were 788 android projects available in Google code repository from which 724 defined at least 1 issue and 183 projects 
had at least one issue including a stack trace. To identify the Android projects on Googlecode we used the search facilities provided looking for 
projects mentioning "Android" on the name of the project or its description. The repository page of each project was inspected, and from this set
142 corresponded to Android applications.

To mine the information available on such Google code projects we implemented a web crawler which navegated on each project and issue page
and extracted the content of each issue to store on relational database. \footnote{Google code previoulsly provided a webservice to extract information about each project, but this service was removed in what Google called "clean-up action" which happened in June 2013.}

@@

Based on the assumption that regardless the issue type, every exception stack
trace contains relevant information concerning the exception structure of the
projects analyzed, we opted for not restricting the analysis only on defect
issues.\footnote{We conducted the same analysis on the defect issues and the top
exceptions are similar to the ones mentioned on defect issues. Due to space
limitation we limit to present the general analysis here. More detailed analysis
on the defect issues can be found at:
\url{www.dimap.ufrn.br/~roberta/android_repo_mining}}

@@

After distilling the information available on the XXXX  stack traces, extracted from the 
issues and issues comments of XXXX Android projects on GitHub and Googlecode,  
we found XXXX  distinct exceptions reported as the root causes of stack traces.
Table \ref{tab:toptenandroid} illustrates the top 10 root couses found in the study ranked by the number of distinct
project repositories on which they were reported.

\begin{table*}
  \centering
  \begin{tabular}{rcccccccc}
    \hline
    \bfseries{Root Exception} &  \multicolumn{2}{c}{\bfseries{Projects}} &  \multicolumn{2}{c}{\bfseries{Occurrences}} & \textsf{Android} & \textsf{Libcore} & \textsf{App} & \textsf{Lib} \\
    & \bfseries{\#} &  \bfseries{\%} & \bfseries{\# } & \bfseries{\% } &&&&\\
    \hline
java.lang.NullPointerException            & 254 & 52,70\%  & 1225 & 30,08\% & 473 & 18 & 595 & 137 \\
java.lang.IllegalStateException          & 99  & 20,54\%  & 234  & 5,75\%  & 165 & 12 & 36  & 20  \\
java.lang.IllegalArgumentException          & 94  & 19,50\% & 255  & 6,26\% & 146 & 6  & 64  & 39   \\
java.lang.RuntimeException                & 91  & 18,88\% & 232  & 5,70\%  & 167 & 1  & 47  & 17   \\
java.lang.OutOfMemoryError                 & 56  & 11,62\% & 180  & 4,42\% & 121 & 15 & 17  & 23   \\
java.lang.NoClassDefFoundError           & 52  & 10,79\%  & 73   & 1,79\%  & 9   & 0  & 37  & 26   \\
java.lang.ClassCastException             & 49  & 10,17\%  & 94   & 2,31\%  & 43  & 0  & 40  & 11   \\
java.lang.IndexOutOfBoundsException       & 47  & 9,75\%   & 127  & 3,12\% & 47  & 0  & 71  & 8    \\
java.lang.NoSuchMethodError               & 40  & 8,30\%  & 57   & 1,40\%  & 9   & 0  & 39  & 9   \\
java.util.ConcurrentModificationException & 38  & 7,88\% & 54   & 1,33\%   & 5   & 0  & 43  & 6   \\

Table \ref{tab:toptenandroid} illustrate the number of times each exception was reported as 
the root cause of a stack trace, the number of distinct projects reporting such causes on stack traces, 
and the percentages of such in relation to all reported exceptions and analyzed projects respectaly.
The table also presents the whether the method responsible for signaling the exceptions were
part of Android, the aplication itself, libcore, or a third-party library as presented in Table  \ref{tab:signalers}  \footnote{9 stack traces contained only elements of JVM and were discarded from the analysis}.

\begin{table}
  \centering
  \begin{tabular}{rp{29em}}
    \hline
    \bfseries{Signaler} & \bfseries{Description} \\
    \hline
    \bfseries{android} & If the exception is thrown by a method defined in Android Platform or OS, or in a JDK library used by them.\\
    \bfseries{app}     & If the exception is thrown by an application method or in a  DK library used by it.\\
    \bfseries{libcore} & If the exception is thrown by one of the core libraries reused by Android (i.e., org.apache.harmony, org.w3c.dom, sun.misc, org.apache.http, org.json, org.xml). \\
    \bfseries{lib}     & If the exception is thrown by a method that was not defined by any of the elements above.\\
    \hline
  \end{tabular}
  \caption{Sources of exceptions in Android}
  \label{tab:signalers}
\end{table}

For each exception we could discover the number of times it occurred on exception stack traces,
the number of distinct projects on which such stacks were defined (and consequently its
popularity) and the of times it was signaled by different signaler types. We
could calculate the popularity of each root cause, considering the number of
distinct projects they were found and the whole set of projects analyzed with at
least one stack trace (482). Table~\ref{tab:toptenandroid} 
%and Figure~\ref{fig:androidsignaler}
presents the mined data.

----------

@@

----------

@@

TODO: SEPARAR JAVA DE ANDROID

\begin{table}
  \centering
  \begin{tabular}{|p{2cm}| p{5cm}|}
    \hline
    \bfseries{Category} & \bfseries{Description} \\
    \hline
      Programming logic &  Exceptions thrown by the JVM when the 
evaluation of an expression violates the normal semantics of language (e.g., 
out-of-bounds array index, division-by-zero, access to a null reference). \\ \hline
      Resources (IO, Memory, Energy)  & Exception releated to the manipulation or restriction of resources (e.g., file, network, memory) \\ \hline
      Security                               & Exception releated to security issues (e.g. password validation, criptography) \\ \hline
      Concurrency                            & Exceptions related to multi-threaded programming \\ \hline
      Backward compatibility                 &  A list of exceptions related to backward compatibility - a complete list can be found here [ref]  \\ \hline
      Reflection                             & Exceptions related to Java reflection library  \\ \hline
      Specific                & Exceptions created for specific applications, frameworks or libraries \\ \hline
      General (Exception, Error, Runtime)    & The top level exceptions in Java    \\ \hline
    \hline
  \end{tabular}
  \caption{Description of categories.}
  \label{tab:categories}
\end{table}

@@

\begin{table}
  \centering
  \begin{tabular}{lrr}
    \hline
    \bfseries{Category} &  \bfseries{ Ocurrences (\%)} \\
    \hline
      Programming logic (java.lang and util) & 2235 (55,64\%)\\
      Resources (IO)                         &  727 (18,10\%) \\
      Security                               &  165 (4,11\%)\\
      Concurrency                            & 116  (2,89\%)\\
      Backward compatibility                 &  219 (5,45\%) \\
      Reflection                             &  91 (2,27\%)\\
      Specific (GUI,FRAMEWORK)               &  197 (4,90\%)\\
      General (Error, Exception, Runtime)    &  267 (6,65\%)\\
    \hline
  \end{tabular}
  \caption{Characterization of the 100 root causes.}
  \label{tab:tophundrend}
\end{table}

@@

@@

To answer this question we: (i) distilled the information available on stack traces (i.e., exception signalers, root causes, exception wrappings); 
(ii) mined the exception types related to each exception reported on stack traces using source code and byte code analyses; and finally checked 
this information against the compiled best practices. Doing so, we mined some characteristics of exception stack traces that points to best practices violations.

 Table~\ref{tab:typeroottab} presents the types of root causes of all stack traces reported on the analyzed issues. As we can see 
from Table ~\ref{tab:typeroottab},  most of the reported exceptions were signaled by Android and libraries reused by Apps were runtime exceptions. 
On the other hand, we could observe that most of the reported exceptions signaled by the set of 
libraries reused by Android platform (i.e., org.apache.harmony,
org.w3c.dom, sun.misc, org.apache.http, org.json, org.xml, javax -- referred in
this work as libcore) were checked exceptions. It is pointed as a good practice
for libraries (see Section~\ref{sec:best}) because by using checked exceptions
libraries can define a precise exception interface to its clients. Such
libraries are heavily used in several projects, and such precise exception
interface may be related to the library's maturity.

\begin{table}
\centering
\begin{tabular}{lccccc}
    \hline
    \bfseries{Root Cause Type} & \bfseries{Android} & \bfseries{Libcore} & \bfseries{App} & \bfseries{Lib}  & \bfseries{All}\\
    \hline

Runtime	&	1075	&	56	&	1446	&	374	&	2951\\
Error	&	144	&	34	&	168	&	105	&		451	\\
Checked	&	110	&	230	&	139	&	145	&	624	\\
Throwable	&	0	&	0	&	1	&	0	&	1	\\
Undefined	&	1	&	0	&	6	&	8	&	15	\\
    \hline
All		& 1330	&	320	&	1760	&	632	&	4042	\\
    \hline
  \end{tabular}
\caption{Types of root exceptions.}
  \label{tab:typeroottab}
\end{table}

%\begin{figure}
%\centering
%\includegraphics[width=\hsize]{chart_exceptiontypes.pdf}
%\includegraphics[width=\hsize]{exception_types.pdf}
%\caption{Distribution of Checked and Unchedked exceptions.}
%\label{fig:typeroot}
%\end{figure}

% REMOVED BECAUSE WE DO NOT HAVE SPACE TO DISCUSS IT, AND IT IS NOT WELL
% CONNECTED WITH PREVIOUS.
%\subsection{Uncaught Checked Exceptions}
%Although most of the  reported exceptions were runtime, we could also find checked
%exceptions as the root causes of stack traces reported on issues. A checked exception
%can remain uncaught in two circumstances: if all methods on the execution trace
%on which this exception flows explicitly specifies the exception, or if the
%checked exception is wrapped in a runtime and than re-thrown. Since checked exceptions 
%should be used to represent recoverable conditions, ignoring a
%checked exception, is considered a bad practice. As mentioned before, when the developer is confronted
%with a checked exception, the designer of the API is telling him to handle the
%exceptional condition. To better understand what was causing checked exceptions
%to scape from being handled we investigated the kinds of wrappings that happened
%on stacks the stacks.  Next sections presents the wrappings found and discuss
%about them.

\subsection{What can the Exception Wrappings tell us?}

Table~\ref{tab:wrappingandroid} presents the wrappings found in this study for all
stack traces found in Android repositories. As we can see, some of the checked
exceptions were indeed wrapped in runtime exceptions or even errors, while most
of them were not wrapped along the stack trace.

\begin{table*}
  \centering
  \begin{tabular}{rccccccc}
    \hline
    \bfseries{Wrapper Type}  &  \bfseries{Root Cause Type} &  \bfseries{\#Projects}  &  \bfseries{\#Occurrences} & \textsf{Android} & \textsf{Libcore} & \textsf{App} & \textsf{Lib}  \\
    \hline
